---
title: "Untitled"
author: "Justin DeCross"
date: "July 25, 2018"
output: html_document
---
### Reading HDF5
#### HDF5
 * Used for storing large data sets
 * Supports storing a range of data types
 * Hierarchical data format
 * groups containing 0 or more data sets and metadata
	+ Have a group header with group name and list of attributes
	+ Have a group symbol table with a list of objects in group
 * datasets multidimensional array of data elements with metadata
	+ Have a header with name, data type, data space, and storage layout
	+ Have a data array with the data
	
#### R HDF5 package

```{r}

source("https://bioconductor.org/biocLite.R")
biocLite("rhdf5")

library(rhdf5)
created  <-  h5createFile("example6.h5")
created
```

 * This will install packages from bioconductor, primary use for genomics but also have good "big data" packages
 * Can be used to interface with HDF5
 
#### Create groups

```{r}
library(rhdf5)
created = h5createGroup("example6.h5","foo")
created = h5createGroup("example6.h5","baa")
created = h5createGroup("example6.h5","foo/foobaa")
h5ls("example6.h5")
```

#### Write to groups

```{r}
A = matrix(1:10, nr=5,nc=2)
h5write(A,"example6.h5","foo/A")
B = array(seq(0.1, 2.0, by= 0.1), dim=c(5,2,2))
attr(B,"scale") <- "liter"
h5write(B,"example6.h5","foo/foobaa/B")
h5ls("example6.h5")
```

#### Write a data set

```{r}
library(rhdf5)
df = data.frame(1L:5L, seq(0,1, length.out=5),
	c("ab","cde","fghi","a","s"), stringsAsFactors=FALSE)
h5write(df,"example6.h5","df")
h5ls("example6.h5")
```

#### Reading data

```{r}
readA = h5read("example6.h5","foo/A")
readB = h5read("example6.h5","foo/foobaa/B")
readds = h5read("example6.h5","df")
readA
```

#### Writing and reading chunks

```{r}
h5write(c(12, 13, 14),"example6.h5","foo/A", index=list(1:3, 1))
h5read("example6.h5", "foo/A")
```

#### Notes and further resources
 * HDF5 can be used to optimize reading/writing from disk in R
